---
title: "Week 2 / Day 2 / Homework"
output: html_notebook
---

# Load Libraries
```{r}
library(tidyverse)
```

# MVP

## Q1.
Load the code_clan_tweets.csv data. Find the number of rows, columns, and list all the variable names.

```{r}
code_clan_tweets <- read_csv("data/code_clan_tweets.csv")
```

```{r}
glimpse(code_clan_tweets)
```

## Q2.
Find the total number of favourites (stored in favorite_count) that CodeClan tweets have got. Don’t add a tweet’s favorite_count to the total when the tweet was quoted (see the is_quote variable).

```{r}
code_clan_tweets %>% 
  filter(is_quote == FALSE) %>% 
  summarise(total_number_of_favourites = sum(favorite_count))
```

## Q3.
Summarise the mean number of retweets (retweet_count), split by the type of platform that was used to tweet (source). Again, omit quoted tweets from the mean.

```{r}
code_clan_tweets %>% 
  filter(is_quote == FALSE) %>% 
  group_by(source) %>% 
  summarise(mean_retweets = mean(retweet_count)) %>% 
  arrange(desc(mean_retweets))
```

## Q4.
Count the total number of likes (i.e. total of favorite_count), split by media type, and arrange them from most likes to least. Assume that any tweets without a listed media type are of type “text”.

```{r}
code_clan_tweets %>% 
  mutate(media_type = replace_na(media_type, "text")) %>% 
  group_by(media_type) %>% 
  summarise(total_likes = sum(favorite_count)) %>% 
  arrange(desc(total_likes))
```
## Q5.
Find the mean number of characters that a CodeClan tweet contains. You can either find a way to count the text in the text variable, or use the variable display_text_width which contains a count. Checking both would be a good idea if possible to see if they match (but it’s not essential).

I tried this using the `str_length()` function and using the `display_text_width` values; however, different values were returned. Looking at a specific tweet `tweet_id` 374, `str_length()` returns 80 and `display_text_width` shows 100. Using Microsoft Word and online word counters the count is shown as 80; however, I haven't been able to identify why the `display_text_width` may be different.

```{r}
code_clan_tweets %>% 
  summarise(avg_character_number = mean(str_length(text)))
```

```{r}
code_clan_tweets %>% 
  summarise(avg_character_number = mean(display_text_width))
```

```{r}
code_clan_tweets %>% 
  filter(tweet_id == 347) %>% 
  mutate(count = str_length(text))
```

## Q6.
The code_clan_info.csv data file contains status_url along with other info. Load this in, and join it to the code_clan_tweets tibble, so that you have a status_url for each tweet. Decide which variable to join the tibbles on.

```{r}
code_clan_info <- read_csv("data/code_clan_info.csv", 
                           col_select = c("tweet_id", "status_url"))
```

```{r}
glimpse(code_clan_info)
```

```{r}
code_clan_tweets_joined <- code_clan_info %>%
  inner_join(code_clan_tweets, code_clan_info, by = "tweet_id")
```


## Q7.
From your new joined data, create a new tibble codeclan_hashtags containing only tweet_id and hashtags, then convert hashtags into lowercase for analysis. Keep only those tweets with hashtags.

```{r}
codeclan_hashtags <- code_clan_tweets_joined %>% 
  select(tweet_id, hashtags) %>% 
  mutate(hashtags = str_to_lower(hashtags)) %>% 
  drop_na() 
```

# Extension

## Q8.
Some tweets have more than one hashtag, denoted by the c( symbols at the start of the string. Use the str_sub() and str_detect() functions and your codeclan_hashtags tibble from above to find all the cases in which the hashtag string begins with charactersc(.

```{r}
pattern <- "c\\("

codeclan_hashtags %>% 
  mutate(hashtags_start = str_sub(hashtags, start = 1, end = 2),
         more_than_one_hashtag = str_detect(hashtags_start, pattern))

```
