---
title: "Week 6 / Day 2"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(e1071)
```

## MVP

### Q1.
Load the data, clean_names() and explore it.

```{r}
phones <- read_csv("data/20190928-items.csv") %>% clean_names()
```

```{r}
glimpse(phones)
head(phones)
```

### Q2.
Find the brand with the highest numbers of phones in the dataset.

```{r}

phones %>%
  group_by(brand) %>% 
  summarise(total_brand_reviews = sum(total_reviews)) %>% 
  arrange(desc(total_brand_reviews))

```

### Q3.
For your top brand, plot the distribution of phone ratings as a probability density, overlaying a fitted normal distribution. Do you think the normal distribution provides a good model of these ratings?

Interpretation:
Based on the plot below where a fitted normal distribution has been plotted over our probability density data, I would be leaning towards normal distribution not being a good model of these ratings as it has a longer left tail and looks as though it may be negatively skewed, 


```{r}

samsung_rating_stats <- phones %>% 
  filter(brand == "Samsung") %>% 
  summarise(num = n(),
            mean = mean(rating),
            sd = sd(rating))

samsung_rating_stats

```

```{r}

phones %>%
  filter(brand == "Samsung") %>% 
  ggplot() +
  geom_histogram(aes(x = rating, y = after_stat(density)), colour = "white") +
  stat_function(
    fun = dnorm,
    args = list(
      mean = samsung_rating_stats$mean,
      sd = samsung_rating_stats$sd
    ),
    colour = "red"
  )

```

### Q4 & Extension
We can be more quantitative in comparing the distribution of top brand ratings with a normal distribution. You calculated the mean() and sd() of your distribution in an earlier question. Use these values to calculate the proportions of ratings within one-, two- and three standard deviations of the mean. Compare these proportions with those you expect for a normal distribution.

Interpretation: 
- Considering the proportion of ratings that fall within 1SD, 2SD and 3SD when compared to what we would expect from a normal distribution. A higher proportion of ratings are within 1SD than we would expect; however, a smaller number are within both 2SD and 3SD. This suggests that a larger proportion than expected in a normal distribution is very close to the mean; however, the overall distribution is more widely spread.

- Considering the QQPLot and QQline, this would suggest that this data does not fit a normal distribution. If it were a normal distribution we would expect the plot points to more closely resemble a straight line and more closely align with the QQline.

- Considering the Skewness of the data -0.9349111, it is moderately negatively skewed, which would suggest the data does not fit a normal distribution.

- The Z Rating calculations show that there are 8 observations which could be considered outliers.

What we'd expect for a normal distribution:

```{r}

100 * (pnorm(q = 1) - pnorm(q = -1))
100 * (pnorm(q = 2) - pnorm(q = -2))
100 * (pnorm(q = 3) - pnorm(q = -3))

```

For Samsung data:

```{r}

samsung_ratings <- phones %>% 
  filter(brand == "Samsung")

# Proportion Within 1 SD

samsung_ratings %>% 
  filter(rating >= samsung_rating_stats$mean - samsung_rating_stats$sd) %>%
  filter(rating <= samsung_rating_stats$mean + samsung_rating_stats$sd) %>%
  summarise(prop_within_1sd = n() / nrow(samsung_ratings))

```

```{r}

# Proportion Within 2 SD

samsung_ratings %>% 
  filter(rating >= samsung_rating_stats$mean - (2 * samsung_rating_stats$sd)) %>%
  filter(rating <= samsung_rating_stats$mean + (2 * samsung_rating_stats$sd)) %>%
  summarise(prop_within_2sd = n() / nrow(samsung_ratings))

```

```{r}

# Proportion Within 3 SD

samsung_ratings %>% 
  filter(rating >= samsung_rating_stats$mean - (3 * samsung_rating_stats$sd)) %>%
  filter(rating <= samsung_rating_stats$mean + (3 * samsung_rating_stats$sd)) %>%
  summarise(prop_within_3sd = n() / nrow(samsung_ratings))

```

                        Expected | Samsung Data    
Proportion Within 1SD   68.26895 | 79.34509 (0.7934509	* 100)
Proportion Within 2SD   95.44997 | 91.93955 (0.9193955	* 100)
Proportion Within 3SD   99.73002 | 97.98489 (0.9798489	* 100)

```{r}

qqnorm(samsung_ratings$rating)
qqline(samsung_ratings$rating, col = "red")

```

```{r}

samsung_ratings %>% 
  ggplot() +
  geom_boxplot(aes(x = rating))

```

```{r}

samsung_ratings %>% 
  summarise(skewness = skewness(rating, type = 1))

```

```{r}

samsung_ratings_scaled <- samsung_ratings %>% 
  mutate(z_rating = scale(rating))

samsung_ratings_scaled %>% 
  filter(!between(z_rating, left = -3, right = 3))

```
