---
title: "Day 2 Homework"
output: html_notebook
---

# Setup

```{r}
# Load Tidyverse
library(tidyverse)
# Load CodeClanData Library
library(CodeClanData)
# View drinks_content to check that it has loaded successfully.
drinks_content
```

# Question 1.

Load in the dataset and have a look at it. What are the dimensions, variable types, variable names, etc.?

### What are the dimensions?

```{r}
dim(drinks_content)

# The data has 242 rows/observations and 18 columns/variables.
```

### What are the variable types?

```{r}
glimpse(drinks_content)

# The data is a mixture of character and numeric data types (dbl = numbers which can contain decimal values.)
```

### What are the variable names?

```{r}
names(drinks_content)
```

# Question 2.

Let's first investigate the calories of different drinks. Select the variables Beverage_category, Beverage, Beverage prep and Calories from your data. Since we are interested in the calorie content, check if there are any NA values in the data, and drop them if there are.

```{r}
# Select the required variables.
selected_drinks <- drinks_content %>% 
  select(Beverage_category:Calories)
```

```{r}
# Check if there any any NA values across all observations and variables.
sum(is.na(selected_drinks))
```

```{r}
# It doesn't look like there were any NA values, but I think this might work for removing any observations which had an NA value in any variable.
selected_drinks_minus_na <- selected_drinks %>% 
  drop_na()
```

# Question 3. Create a new variable (column) called calorie_diff, which stores the difference between 135 calories (135 calories = 10 cubes of sugar!) and the calories in each drink. (hint: you'll want to subtract 135 from the calories to see which drink have more or less than 10 cups of sugar).

```{r}
selected_drinks %>%
  mutate(calorie_diff = Calories - 135)
```

# Question 4.

Summarise the mean number of calories in each beverage_category. Which 3 category of drinks have the most calories? Which 3 drinks have the least? Write a small summary of your findings.

### Summarise the mean number of calories in each beverage_category

```{r}
selected_drinks %>%
  group_by(Beverage_category) %>% 
  summarise(mean_calories = mean(Calories))
```

### Which 3 category of drinks have the most calories?

```{r}
selected_drinks %>%
  group_by(Beverage_category) %>% 
  summarise(mean_calories = mean(Calories)) %>% 
  arrange(desc(mean_calories)) %>% 
  head(3)
```

### Which 3 drinks have the least?

```{r}
selected_drinks %>%
  arrange(Calories) %>% 
  head(3)
```

### Write a small summary of your findings.

Smoothies are the beverage category with th most average calories, followed by Frappuccino's and Signature Espresso's.

Tall, Grande and Venti Tazo Teas are drinks which all have 0 Calories. However, a flaw in viewing the data this way is that it misses that a Short Tazo Tea also has 0 Calories as confirmed below.

```{r}
selected_drinks %>%
  filter(Calories == 0)
```

# Question 5.

Let's look at this a different way. What is the average number of calories in each Beverage_prep type?

```{r}
selected_drinks %>%
  group_by(Beverage_prep) %>% 
  summarise(mean_calories = mean(Calories))
```

# Question 6.

Which Beverage_prep type contains more than the average calories of all drinks?

```{r}
# Calculate the average calories across all drinks.
average_calories <- selected_drinks %>% 
  summarise(average_calories = mean(Calories)) %>% 
  pull()
```

```{r}
# Filter the mean_calories figure for each Beverage Prep using the calculated average_calories across all drinks.
selected_drinks %>% 
  group_by(Beverage_prep) %>% 
  summarise(prep_mean_calories = mean(Calories)) %>% 
  filter(prep_mean_calories > average_calories)
```

# Question 7.

Which is the best type of coffee drink (and by 'drink' we mean the combination of Beverage_category, Beverage and Beverage_prep) to get if you're worried about consuming too many calories?

```{r}
drinks_content %>% 
  filter(Beverage_category == "Coffee") %>% 
  arrange(Calories) %>% 
  select(Beverage_category, Beverage, Beverage_prep, Calories) %>% 
  head(1)
```
