---
title: "Week 10, Day 1"
output: html_notebook
---

```{r}
library(tidyverse)
library(broom)
library(ggfortify)
```

# MVP

## 1.
Load the data into a dataframe project

```{r}
project_management <- read_csv("data/project_management.csv")
```

## 2.
Plot the data, taking estimated_length as the independent variable and actual_length as the dependent variable.

Independent Variable / Predictor / X / estimated_length
Dependent Variable / Response / Y / actual_length

```{r}
project_management %>% 
  ggplot() +
  geom_point(aes(x = estimated_length, y = actual_length))
```

## 3.
Calculate the correlation coefficient of estimated_length and actual_length and interpret the value you obtain.

```{r}
project_management %>%
  summarise(correlation_coefficient = cor(estimated_length, actual_length))
```

The correlation coefficient of estimated_length and actual_length is returned as 0.8046597. This suggests that their is a very strong correlation between estimated_length and actual_length.

## 4.
Perform a simple linear regression using actual_length as the dependent variable, and estimated_length as the independent variable. Save the model object to a variable.

```{r}
model <- lm(formula = actual_length ~ estimated_length, 
             data = project_management)

model 
```

## 5.
Interpret the regression coefficient of estimated_length (i.e. slope, gradient) you obtain from the model. How do you interpret the r2 value reported by the model?

The coefficient of estimated_length is returned as 1.223. This means that for every unit increase of estimated_length the actual_length is expected to increase by 1.223 units.

The r-squared value returned is 0.6474772. We can therefor say that our simple linear model says that 65% of the variation in actual_duration can be explained by the estimated_duration.

```{r}
glance(model)
```

## 6.
Is the relationship statistically significant? Remember, to assess this you need to check the p
-value of the regression coefficient (or slope/gradient). But you should first check the regression diagnostic plots to see if the p-value will be reliable (don’t worry about any outlier points you see in the diagnostic plots, we’ll return to them in the extension).

- Residuals vs. Fitted
There doesn't appear to be any strong pattern here, which supports the reliability of the p-value.

- Normal QQ
There distribution of the residuals appears to be mostly normal, with the exception of a couple of outliers, supporting the reliability of the p-value.

- Scale Location
The spread of residuals appears to be fairly consistent across all fitted values, supporting reliability of the p-value.

The p-value returned is very low, suggesting that the relationship is statistically significant.


```{r}
autoplot(model)
```



