---
title: "Week 11, Day 5"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(GGally)
library(modelr)
library(broom)
library(pROC)
library(caret)
```

# MVP

```{r}
juice <- read_csv("data/orange_juice.csv") %>% clean_names()
```

```{r}
glimpse(juice)
```

```{r}
summary(juice)
```

## Data Preperation

### 1. Replace `purchase` with logical `purchase_mm`

```{r}
juice_clean <- juice %>% 
  mutate(purchase_mm = if_else(purchase == "MM", TRUE, FALSE), .after = "purchase")

juice_clean <- juice_clean %>% 
  mutate(purchase_mm = as.factor(purchase_mm)) %>% 
  select(-purchase)
```

### 2. Update `store_id`, `special_ch`, `special_mm`, `store7` & `store` To Factors

```{r}

juice_clean <- juice_clean %>%
  mutate(store_id = as.factor(store_id),
         special_ch = as.factor(special_ch),
         special_mm = as.factor(special_mm),
         store7 = as.factor(store7),
         store = as.factor(store))

```

### 3. Decide On Management of `weekof_purchase` Variable

For this task, I am going to assume that week of purchase would correlate in some way to time periods where discounts where offered on either brand. I will also assume that it is the discount that is the predicting factor, not the week.

As data on data on discounts at the time of purchase is available through other variables, for now, I will remove the `weekof_purchase` variable.

```{r}

juice_clean <- juice_clean %>%
  select(-weekof_purchase)

```


### 4. Wrangle `store_id`, `store7` & `store` Variables.

Based on the counts `store` and `store_id` the only difference is that `store` lists a store 0 with 356 observations and `store_id` lists a store 7 with 356 observations. These therefor look to be the same store. As there is an additonal `store7` variable, I will assume that 7 is the correct variable and drop the `store` and `store7` variables.

```{r}
juice %>% 
  count(store)
```

```{r}
juice %>% 
  count(store_id)
```

```{r}

juice_clean <- juice_clean %>%
  select(-c(store, store7))

```

### 5. Check for Aliases Variables

Findings...

`sale_price_mm` aliased with `price_mm` and `disc_mm` / `sale_price_mm` will be removed.
`sale_price_ch` aliased with `price_ch` and `disc_ch` / `sale_price_ch` will be removed.
`price_diff` aliased with `price_ch`, `price_mm`, `disc_ch` & `disc_mm` / `price_diff` will be removed.
`list_price_diff` aliased with `price_ch` & `price_mm` / `list_price_diff` will be removed.

```{r}

juice_clean <- juice_clean %>%
  select(-c(sale_price_mm, sale_price_ch, -price_diff, -list_price_diff))

```

```{r}
alias(purchase_mm ~ ., data = juice_clean)
```

### 6. Check for Strongly Correlated Potential Predictors

There are still a significant number of variables that relate to price.

- `price_ch`
- `price_mm`
- `disc_ch`
- `disc_mm`
- `special_ch`
- `special_mm`
- `price_diff`
- `pct_disc_mm`
- `pct_disc_ch`
- `list_price_diff`

Having checked to see if any of these are strongly correlated to one another and whether any should be dropped:

- `pct_disc_mm` and `disc_mm` have a very strong positive correlation of 0.999 which is statistically significant.
- `pct_disc_ch` and `disc_ch` have a very strong positive correlation of 0.999 which is statistically significant.

This is to be expected and for the purposes of this exercise I will drop `pct_disc_mm` and `pct_disc_ch`.

```{r message=FALSE}

juice_clean %>% 
  select(-c(purchase_mm, store_id, loyal_ch)) %>% 
  ggpairs(progress = FALSE)

```

```{r}

juice_clean <- juice_clean %>%
  select(-c(pct_disc_mm, pct_disc_ch))

```

## Explore Potential Predictors

To make it easier to initially explore potential predictor candidates, I have separated the variables into 3 groups.

Group 1 / Price: `price_ch`, `price_mm`, `disc_ch`, `disc_mm`, `special_ch`, `special_mm`, `price_diff`, `list_price_diff`
Group 2 / Store: `store_id`
Group 3 / Customer Attributes: `loyal_ch`

### 1. Explore Group 1

The below look like the most promising potential predictors.

- `disc_ch`
- `disc_mm`
- `price_diff`
- `list_price_diff`

```{r message=FALSE}
juice_clean %>% 
  select(-c(store_id, loyal_ch)) %>% 
  ggpairs(progress = FALSE)
```

### 2. Explore Group 2

There does also appear to be some correlation with `store_id`.

```{r message=FALSE}
juice_clean %>% 
  select(purchase_mm, store_id) %>% 
  ggpairs(progress = FALSE)
```

### 3. Explore Group 3

There does also appear to be some correlation with `loyal_ch`.

```{r}
juice_clean %>% 
  select(purchase_mm, loyal_ch) %>% 
  ggpairs(progress = FALSE)
```

## Check Predictors for Statistical Significance

### `disc_ch`

`disc_ch` appears to be a statistically significant predictor.

```{r}
# From Group 1 / `disc_ch` as predictor

juice_discch_model <- glm(
  purchase_mm ~ disc_ch,
  data = juice_clean,
  family = binomial(link = "logit")
)

summary(juice_discch_model)
```

### `disc_mm`

`disc_mm` appears to be a statistically significant predictor.

```{r}
# From Group 1 / `disc_mm` as predictor

juice_discmm_model <- glm(
  purchase_mm ~ disc_mm,
  data = juice_clean,
  family = binomial(link = "logit")
)

summary(juice_discmm_model)
```

### `price_diff`

`price_diff` appears to be a statistically significant predictor.

```{r}
# From Group 1 / `price_diff` as predictor

juice_pricediff_model <- glm(
  purchase_mm ~ price_diff,
  data = juice_clean,
  family = binomial(link = "logit")
)

summary(juice_pricediff_model)
```

### `list_price_diff`

`list_price_diff` appears to be a statistically significant predictor.

```{r}
# From Group 1 / `loyalch` as predictor

juice_listpricediff_model <- glm(
  purchase_mm ~ list_price_diff,
  data = juice_clean,
  family = binomial(link = "logit")
)

summary(juice_listpricediff_model)
```

### `loyal_ch`

`loyal_ch` appears to be a statistically significant predictor.

```{r}

# From Group 2 / `loyal_ch` as predictor

juice_loyalch_model <- glm(
  purchase_mm ~ loyal_ch,
  data = juice_clean,
  family = binomial(link = "logit")
)

summary(juice_loyalch_model)
```

### `store_id`

Only some `store_id` values appear to be statistically significant predictors.

```{r}

# From Group 3 / `store_id` as predictor

juice_storeid_model <- glm(
  purchase_mm ~ store_id,
  data = juice_clean,
  family = binomial(link = "logit")
)

summary(juice_storeid_model)
```

### Findings

loyal_ch / Statistically Significant / B1 Coefficient -6.0948
disc_ch / Statistically Significant / B1 Coefficient -3.9819
list_price_diff / Statistically Significant / B1 Coefficient -3.1438
price_diff / Statistically Significant / B1 Coefficient -2.16873
disc_mm / Statistically Significant / B1 Coefficient 1.57365
store_id / IDs 3, 4 & 7 Statistically Significant.

## Add Predictions And Calculate AUC For Different Predictor Combinations

I will add predictions in the order of B1 Coefficient size and record AUC values.

### 1x Predictor

```{r}
juice_model_1_pred <- glm(
  purchase_mm ~ loyal_ch,
  data = juice_clean,
  family = binomial(link = "logit")
)

juice_model_1_summary <- clean_names(tidy(juice_model_1_pred))

juice_model_1_summary

juice_preds <- juice_clean %>% 
  add_predictions(juice_model_1_pred,
                  type = "response")

roc_obj_juice_1_pred <- juice_preds %>% 
  roc(response = purchase_mm,
      predictor = pred)

auc(roc_obj_juice_1_pred)
```

### 2x Predictor

```{r}

juice_model_2_pred <- glm(
  purchase_mm ~ loyal_ch + disc_ch,
  data = juice_clean,
  family = binomial(link = "logit")
)

juice_model_2_summary <- clean_names(tidy(juice_model_2_pred))

juice_model_2_summary

juice_preds <- juice_clean %>% 
  add_predictions(juice_model_2_pred,
                  type = "response")

roc_obj_juice_2_pred <- juice_preds %>% 
  roc(response = purchase_mm,
      predictor = pred)

auc(roc_obj_juice_2_pred)
```

### 3x Predictor

```{r}

juice_model_3_pred <- glm(
  purchase_mm ~ loyal_ch + disc_ch + list_price_diff,
  data = juice_clean,
  family = binomial(link = "logit")
)

juice_model_3_summary <- clean_names(tidy(juice_model_3_pred))

juice_model_3_summary

juice_preds <- juice_clean %>% 
  add_predictions(juice_model_3_pred,
                  type = "response")

roc_obj_juice_3_pred <- juice_preds %>% 
  roc(response = purchase_mm,
      predictor = pred)

auc(roc_obj_juice_3_pred)
```

### 4x Predictor

```{r}

juice_model_4_pred <- glm(
  purchase_mm ~ loyal_ch + disc_ch + list_price_diff + price_diff,
  data = juice_clean,
  family = binomial(link = "logit")
)

juice_model_4_summary <- clean_names(tidy(juice_model_4_pred))

juice_model_4_summary

juice_preds <- juice_clean %>% 
  add_predictions(juice_model_4_pred,
                  type = "response")

roc_obj_juice_4_pred <- juice_preds %>% 
  roc(response = purchase_mm,
      predictor = pred)

auc(roc_obj_juice_4_pred)
```

### 5x Predictor A

```{r}
juice_model_5_pred <- glm(
  purchase_mm ~ loyal_ch + disc_ch + list_price_diff + price_diff + disc_mm,
  data = juice_clean,
  family = binomial(link = "logit")
)

juice_model_5_summary <- clean_names(tidy(juice_model_5_pred))

juice_model_5_summary

juice_preds <- juice_clean %>% 
  add_predictions(juice_model_5_pred,
                  type = "response")

roc_obj_juice_5_pred <- juice_preds %>% 
  roc(response = purchase_mm,
      predictor = pred)

auc(roc_obj_juice_5_pred)
```

### 5x Predictor B

```{r}
juice_model_5_pred_b <- glm(
  purchase_mm ~ loyal_ch + disc_ch + list_price_diff + price_diff + store_id,
  data = juice_clean,
  family = binomial(link = "logit")
)

juice_model_5_summary_b <- clean_names(tidy(juice_model_5_pred_b))

juice_preds <- juice_clean %>% 
  add_predictions(juice_model_5_pred_b,
                  type = "response")

roc_obj_juice_5_pred_b <- juice_preds %>% 
  roc(response = purchase_mm,
      predictor = pred)

auc(roc_obj_juice_5_pred_b)
```

### Findings

Model Name / Formula / AUC

juice_model_1_pred / purchase_mm ~ loyal_ch / AUC 0.8808
juice_model_2_pred / purchase_mm ~ loyal_ch + disc_ch / AUC 0.8808
juice_model_3_pred / purchase_mm ~ loyal_ch + disc_ch + list_price_diff / AUC 0.8895
juice_model_4_pred / purchase_mm ~ loyal_ch + disc_ch + list_price_diff + price_diff / AUC 0.8999
juice_model_5_pred / purchase_mm ~ loyal_ch + disc_ch + list_price_diff + price_diff + disc_mm / AUC 0.8999
juice_model_5_pred_b / purchase_mm ~ loyal_ch + disc_ch + list_price_diff + price_diff + store_id / AUC 0.9022

We can see from the above results that the below model returns the highest AUC.

purchase_mm ~ loyal_ch + disc_ch + list_price_diff + price_diff + store_id

### Plot ROC Curve for All 6 Models

We can also plot the ROC Curve for all 6 models to visual the above results. From this, we can see that "5b" does indeed appear to be the best option at this stage.

```{r}
roc_curve <- ggroc(data = list("1" = roc_obj_juice_1_pred,
                               "2" = roc_obj_juice_2_pred,
                               "3" = roc_obj_juice_3_pred,
                               "4" = roc_obj_juice_4_pred,
                               "5" = roc_obj_juice_5_pred,
                               "5b" = roc_obj_juice_5_pred_b),
                   legacy.axes = TRUE)+
  coord_fixed()

roc_curve
```

## Cross Validation

To for over-fit, we will use cross-validation.

### Update TRUE/FALSE To "Yes"/"No" for `purchase_mm`

```{r}


juice_clean_fct <- juice_clean %>%
  mutate(purchase_mm = if_else(purchase_mm == TRUE, "Yes", "No"),
         purchase_mm = as.factor(purchase_mm))

```

### Set Training Specifications

```{r}
train_control <-  trainControl(method = "repeatedcv",
                               number = 5,
                               repeats = 100,
                               savePredictions = TRUE,
                               classProbs = TRUE,
                               summaryFunction = twoClassSummary)
```

### Check Validated AUC/ROC Values

```{r}
juice_model_1_pred_validate <- train(purchase_mm ~ loyal_ch,
               data = juice_clean_fct,
               trControl = train_control,
               method = "glm",
               family = binomial(link = "logit"))

juice_model_1_pred_validate$results$ROC
```

```{r}
juice_model_2_pred_validate <- train(purchase_mm ~ loyal_ch + disc_ch,
               data = juice_clean_fct,
               trControl = train_control,
               method = "glm",
               family = binomial(link = "logit"))

juice_model_2_pred_validate$results$ROC
```

```{r}
juice_model_3_pred_validate <- train(purchase_mm ~ loyal_ch + disc_ch + list_price_diff,
               data = juice_clean_fct,
               trControl = train_control,
               method = "glm",
               family = binomial(link = "logit"))

juice_model_3_pred_validate$results$ROC
```

```{r}
juice_model_4_pred_validate <- train(purchase_mm ~ loyal_ch + disc_ch + list_price_diff + price_diff,
               data = juice_clean_fct,
               trControl = train_control,
               method = "glm",
               family = binomial(link = "logit"))

juice_model_4_pred_validate$results$ROC
```

```{r}
juice_model_5_pred_validate <- train(purchase_mm ~ loyal_ch + disc_ch + list_price_diff + price_diff + disc_mm,
               data = juice_clean_fct,
               trControl = train_control,
               method = "glm",
               family = binomial(link = "logit"))

juice_model_5_pred_validate$results$ROC
```

```{r}
juice_model_5b_pred_validate <- train(purchase_mm ~ loyal_ch + disc_ch + list_price_diff + price_diff + store_id,
               data = juice_clean_fct,
               trControl = train_control,
               method = "glm",
               family = binomial(link = "logit"))

juice_model_5b_pred_validate$results$ROC
```

### Results

juice_model_1_pred
purchase_mm ~ loyal_ch
AUC 0.8808 / Cross Validated AUC 0.8741937
AUC Diff. 0.0066063

juice_model_2_pred
purchase_mm ~ loyal_ch + disc_ch
AUC 0.8808 / Cross Validated AUC 0.8801089
AUC Diff. 0.0006911

juice_model_3_pred
purchase_mm ~ loyal_ch + disc_ch + list_price_diff
AUC 0.8895 / Cross Validated AUC 0.8880731
AUC Diff. 0.0014269

juice_model_4_pred
purchase_mm ~ loyal_ch + disc_ch + list_price_diff + price_diff
AUC 0.8999 / Cross Validated AUC 0.8982
AUC Dif. 0.0017

juice_model_5_pred 
purchase_mm ~ loyal_ch + disc_ch + list_price_diff + price_diff + disc_mm
AUC 0.8999 / Cross Validated AUC 0.8984135
AUC Dif. 0.0014865

juice_model_5_pred_b
purchase_mm ~ loyal_ch + disc_ch + list_price_diff + price_diff + store_id
AUC 0.9022 / Cross Validated AUC 0.8981727
AUC Dif. 0.0040273

### Interpretation

- None of the models have a large decrease in AUC following Cross Validation, suggesting that none of the models are over-fit.

- Although the Cross Validated AUC increases as the number of predictors increase, it could be argued that the increases are small. For example, if we were to round the Cross Validated AUC to 2x decimal places then the model with 4 predictors would have the same Cross Validated AUC as the two models wit 5x predictors (0.90).

- With the above in mind, to balance AUC score with simplicity, I would interpret the below model as being the best predictive classifier in this instance:

juice_model_4_pred
purchase_mm ~ loyal_ch + disc_ch + list_price_diff + price_diff
AUC 0.8999 / Cross Validated AUC 0.8982
AUC Diff. 0.0017