---
title: "NHS Borders"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
  pdf_document: default
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}

# Load Libraries
library(tidyverse)
library(janitor)

# Set Colour Scheme

colour_scheme <- c("#092869", "#0391BF", "#6B077B")

# Read In Data

cancer_incidence <- read_csv("data/cancer_incidence_by_health_board.csv") %>% 
  clean_names()

five_year_summary <- read_csv("data/five_year_cancer_incidence_summary.csv") %>% 
  clean_names()

geography_codes <- read_csv("data/geography_codes_and_labels.csv") %>%
  clean_names()
```


```{r, echo = FALSE, warning = FALSE, message = FALSE}

# cancer_incidence Borders Only

cancer_incidence_borders <- cancer_incidence %>% 
  filter(hb == "S08000016")


cancer_incidence_scotland_borders <- 

# Clean five_year_summary

five_year_summary_borders <- five_year_summary %>% 
  filter(hb == "S08000016")

five_year_summary_borders_incidence <- five_year_summary_borders %>%
  select(id, hb, cancer_site, sex, year | starts_with("incidences"))

five_year_summary_long <- five_year_summary_borders_incidence %>%
  pivot_longer(
    cols = incidences_age_under5:incidences_age85and_over,
    names_to = "age",
    values_to = "incidences"
  ) %>% 
  select(-incidences_all_ages) %>% 
  mutate(age = str_remove(age, "incidences_age"),
         age = str_replace_all(age, "_", ""),
         age = str_replace_all(age, "to", "-"),
         age = case_when(
           age == "under5" ~ "Under 5",
           age == "85andover" ~ "85+",
           TRUE ~ age)) %>% 
  mutate(age = factor(age, levels = c("Under 5", "5-9",
                                      "10-14", "15-19",
                                      "20-24", "25-29",
                                      "30-34", "35-39",
                                      "40-44", "45-49",
                                      "50-54", "55-59",
                                      "60-64", "65-69",
                                      "70-74", "75-79",
                                      "80-84", "85+")))

```

# Overview

Across all ages and genders, the 3 forms of cancer with the highest number of incidences in NHS Borders between 1997-2021 are:

**Cancer Site / Incidences**

1. Non-melanoma skin cancer /	6174
2. Basal cell carcinoma of the skin	/ 4049
3. Breast	/ 2614

> *NB - Basal cell carcinoma of the skin is form form of Non-melanoma skin cancer (NHS Inform, 2022)*

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# cancer_incidence_borders %>% 
#   filter(cancer_site != "All cancer types",
#          sex == "All") %>% 
#   group_by(cancer_site) %>% 
#   summarise(total_incidences = sum(incidences_all_ages)) %>% 
#   arrange(desc(total_incidences)) %>% 
#   head(3)
```

For the purposes of clarity, this report is not intended to be and is by no means an exhaustive overview of cancer incidences and rates within NHS Borders and Scotland as a whole. This report will focus on discussion around 2 cancer types that show the largest incidence numbers in NHS Borders between 1997 and 2021.

1. Non-melanoma skin cancer.
2. Breast cancer.

***

# Breast Cancer

## Context

In the available date, among females in NHS Borders,breast cancer has the highest number of incidences, highest mean crude rate and highest mean European age-standardised rate (EASR).

<div class = "row">
  
<div class = "col-md-6">
**Cancer Site / Incidences**

**1.** Breast / 2598			
**2.** Non-melanoma skin cancer / 2519			
**3.** Basal cell carcinoma of the skin / 1882	
</div>
  
<div class = "col-md-6">
**Cancer Site / Mean Crude Rate**

**1.** Breast	/ 179.9348			
**2.** Non-melanoma skin cancer /	174.2811			
**3.** Basal cell carcinoma of the skin /	130.1987	
</div>

</div>

<div class = "row">  
<div class = "col-md-6">
**Cancer Site / Mean EASR**

**1.** Breast / 161.3640			
**2.** Non-melanoma skin cancer / 150.3996			
**3.** Basal cell carcinoma of the skin / 113.9178			
</div>


</div>

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# cancer_incidence_borders %>% 
#   filter(cancer_site != "All cancer types",
#          sex == "Females") %>% 
#   group_by(cancer_site) %>% 
#   summarise(total_incidences = sum(incidences_all_ages)) %>% 
#   arrange(desc(total_incidences)) %>% 
#   head(3)
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# cancer_incidence_borders %>% 
#   filter(cancer_site != "All cancer types",
#          sex == "Females") %>% 
#   group_by(cancer_site) %>% 
#   summarise(mean_crude_rate = mean(crude_rate)) %>% 
#   arrange(desc(mean_crude_rate)) %>% 
#   head(3)
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# cancer_incidence_borders %>% 
#   filter(cancer_site != "All cancer types",
#          sex == "Females") %>% 
#   group_by(cancer_site) %>% 
#   summarise(mean_easr = mean(easr)) %>% 
#   arrange(desc(mean_easr)) %>% 
#   head(3)
```

## Visualisations

### *[fig. 1]*

```{r, echo = FALSE, warning = FALSE, message = FALSE}

cancer_incidence_borders %>% 
  filter(sex == "Females",
         cancer_site %in% c("All cancer types", "Breast")) %>% 
  ggplot() +
  geom_line(aes(x = year, y = incidences_all_ages, colour = cancer_site),
            size = 1.5) +
  geom_point(aes(x = year, y = incidences_all_ages, colour = cancer_site),
            size = 2) +
  scale_x_continuous(breaks = c(1997, 1999, 2001, 2003, 2005, 2007, 2009,
                                2011, 2013, 2015, 2017, 2019, 2021)) +
  scale_colour_manual(values = colour_scheme, labels = c("All Combined", "Breast")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  ylim(0, 500) +
  labs(
    x = "\n Year",
    y = "Incidences\n",
    title = "All Female Cancer Incidences vs. Female Breast Cancer Incidences",
    subtitle = "NHS Borders: 1997-2021",
    colour = "Cancer Type:") +
  geom_vline(xintercept = 2020, linetype = "dashed") +
  annotate("text", x = 2021, y = 500, label = "COVID-19", size = 2, face = "bold", colour = "red") +
  theme(plot.title = element_text(size = 13, face = "bold"),
        plot.subtitle = element_text(size = 10),
        panel.background = element_rect(fill = "white"),
        panel.grid = element_line(colour = "grey90", 
                                  linetype = "dashed"))

```

**What does this visualisation tell us?**

- That breast cancer incidences in females are driving the trend of all cancer type incidences.
- That there appears to be a historic trend in breast cancer incidences where they peak approximately every 3 years.
- That this trend is less evident in 2020 when we may have expected it, although there is an increase in 2021.

**Why might there be a historic 3 year trend?**

Women who meet screening criteria are invited for breast screening once every 3 years *(NHS National Services Scotland, 2022)*.

**Why might we not see the same peak in 2020 as we may have expected?**

Due to the COVID-19 pandemic, no invites to breast screenings were sent between 30 March 2020 and 3 August 2020 *(Public Health Scotland, 2022)*.

***

### *[fig. 2]*

```{r, echo = FALSE, warning = FALSE, message = FALSE}
five_year_summary_long %>%
  filter(cancer_site == "Breast",
         sex == "Females") %>%
  ggplot() +
  geom_col(aes(x = age, y = incidences),
           fill = "#0391BF") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  labs(
    x = "\n Age",
    y = "Incidences\n",
    title = "Total Female Breast Cancer Incidences by Age",
    subtitle = "NHS Borders: 2017-2021") +
  theme(plot.title = element_text(size = 15, face = "bold"),
        plot.subtitle = element_text(size = 10),
        panel.background = element_rect(fill = "white"),
        panel.grid = element_line(colour = "grey90", 
                                  linetype = "dashed"))
```

**What does this visualisation tell us?**

- The majority of breast cancer incidences in females appear to be between those aged between 50 and 79.

**Why might these age groups see increased incidence numbers?**

- Currently only women between the ages of 50 and 70 are routinely screened *(NHS National Services Scotland, 2022)*.

<style>
div.yellow {background-color:#fcf8c0; border-radius: 5px; padding: 20px;}

</style>
<div class = "yellow">

**NHS Borders Population Projections:**

Females 50+ 2021: 29889

Females 50+ 2041: 31148 (4.21225% increase)

*(National Records of Scotland, 2023)*

</div>

***

## Conclusions / Recommendations

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

- We could reasonably expect to see a continued increase in breast cancer incidences in 2022 as screening appointments catch up following the disruption caused by the COVID-19 pandemic. Resources should be allocated accordingly.

- Consideration should be given, informed by evidence, of expanding screening criteria to lower age groups to determine whether or not the current distribution of incidences by age is due to current screening criteria.

- A projected population increase of ~4% over the 20 years from the end of this data set should be taken into consideration in long term planning.

</div>

***

# Non-melanoma Skin Cancer

## Context


In the available date, among males in NHS Borders, Non-melanoma skin cancer has the highest number of incidences, highest mean crude rate and highest mean European age-standardised rate (EASR).

<div class = "row">
  
<div class = "col-md-6">
**Cancer Site / Incidences**

**1.** Non-melanoma skin cancer	/ 3655			
**2.** Prostate	/ 2253				
**3.** Basal cell carcinoma of the skin	/ 2167	
</div>
  
<div class = "col-md-6">
**Cancer Site / Mean Crude Rate**

**1.** Non-melanoma skin cancer	268.35849			
**2.** Prostate	165.64352			
**3.** Basal cell carcinoma of the skin	159.20984		
</div>

</div>

<div class = "row">  
<div class = "col-md-6">
**Cancer Site / Mean EASR**

**1.** Non-melanoma skin cancer	/ 274.87691			
**2.** Prostate	/ 161.63449				
**3.** Basal cell carcinoma of the skin	/ 156.12175				
</div>


</div>

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# cancer_incidence_borders %>% 
#   filter(cancer_site != "All cancer types",
#          sex == "Male") %>% 
#   group_by(cancer_site) %>% 
#   summarise(total_incidences = sum(incidences_all_ages)) %>% 
#   arrange(desc(total_incidences)) %>% 
#   head(5)
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# cancer_incidence_borders %>% 
#   filter(cancer_site != "All cancer types",
#          sex == "Male") %>% 
#   group_by(cancer_site) %>% 
#   summarise(mean_crude_rate = mean(crude_rate)) %>% 
#   arrange(desc(mean_crude_rate)) %>% 
#   head(5)
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# cancer_incidence_borders %>% 
#   # filter(cancer_site != "All cancer types",
#   #        sex == "Male") %>% 
#   # group_by(cancer_site) %>% 
#   # summarise(mean_easr = mean(easr)) %>% 
#   # arrange(desc(mean_easr)) %>% 
#   # head(5)
```

***

## Visualisations

### *[fig. 3]*

```{r, echo = FALSE, warning = FALSE, message = FALSE}
cancer_incidence_borders %>% 
  filter(sex == "Male",
         cancer_site %in% c("All cancer types", "Non-melanoma skin cancer")) %>% 
  ggplot() +
  geom_line(aes(x = year, y = incidences_all_ages, colour = cancer_site),
            size = 1.5) +
  geom_point(aes(x = year, y = incidences_all_ages, colour = cancer_site),
            size = 2) +
  geom_line(aes(x = year, y = incidences_all_ages, colour = cancer_site),
            size = 1.5) +
  geom_point(aes(x = year, y = incidences_all_ages, colour = cancer_site),
            size = 2) +
  scale_x_continuous(breaks = c(1997, 1999, 2001, 2003, 2005, 2007, 2009,
                                2011, 2013, 2015, 2017, 2019, 2021)) +
  scale_colour_manual(values = colour_scheme, labels = c("All Combined", "Non-melanoma Skin Cancer")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  ylim(0, 500) +
  labs(
    x = "\n Year",
    y = "Incidences\n",
    title = "All Male Cancer Incidences vs. Male Non-melanoma Skin Cancer Incidences",
    subtitle = "NHS Borders: 1997-2021",
    colour = "Cancer Type:") +
  geom_vline(xintercept = 2020, linetype = "dashed") +
  annotate("text", x = 2021, y = 500, 
           label = "COVID-19", 
           size = 2, face = "bold", colour = "red") +
  theme(plot.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 8),
        panel.background = element_rect(fill = "white"),
        panel.grid = element_line(colour = "grey90", 
                                  linetype = "dashed"))
```

**What does this visualisation tell us?**

- That incidences of male non-melanoma skin cancer are generally following the same trend as incides of all types of cancer in men, slowly increasing from 1997 to 2021.

- There is no initially obvious trend as seen with breast cancer incidences and the impact of the COVID-19 Pandemic on incidences is also less clear.

***

### *[fig. 4]*

```{r, echo = FALSE}
five_year_summary_long %>%
  filter(cancer_site == "Non-melanoma skin cancer",
         sex == "Male") %>%
  ggplot() +
  geom_col(aes(x = age, y = incidences),
           fill = "#0391BF") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  labs(
    x = "\n Age",
    y = "Incidences\n",
    title = "Total Male Non-melanoma Skin Cancer Incidences by Age",
    subtitle = "NHS Borders: 2017-2021") +
  theme(plot.title = element_text(size = 15, face = "bold"),
        plot.subtitle = element_text(size = 10),
        panel.background = element_rect(fill = "white"),
        panel.grid = element_line(colour = "grey90", 
                                  linetype = "dashed"))
```

**What does this visualisation tell us?**

- The majority of Non-melanoma skin cancer Incidences in males appear to be between those aged between 70 and 85+.

<style>
div.yellow {background-color:#fcf8c0; border-radius: 5px; padding: 20px;}

</style>
<div class = "yellow">

**NHS Borders Population Projections:**

Males 70+ 2021: 9692

Males 70+ 2041: 13487 (39.156% increase)

*(National Records of Scotland, 2023)*

</div>

***

## Conclusions / Recommendations

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

- Key to effective server precision will be plans to accomdate the projected population increase of ~39.156% over the 20 years from the end of this data set.

</div>

***

# References
- NHS Inform, 2022:
https://www.nhsinform.scot/illnesses-and-conditions/cancer/cancer-types-in-adults/skin-cancer-non-melanoma

- NHS National Services Scotland, 2022: https://www.nss.nhs.scot/specialist-healthcare/screening-programmes/breast-screening/

- National Records of Scotland, 2023:
https://www.nrscotland.gov.uk/statistics-and-data/statistics/statistics-by-theme/population/population-projections/sub-national-population-projections/2018-based/detailed-datasets

- Public Health Scotland, 2022:
https://www.publichealthscotland.scot/media/12843/2022-04-26_breast_screening_report.pdf
