---
title: "Week 10 / Day 2"
output: html_notebook
---

```{r}
library(tidyverse)
library(GGally)
library(ggfortify)

```

# 1.
Load the housing_prices.csv data set and undertake an initial exploration of the data.

```{r}
housing <- read_csv("data/housing.csv")

housing_prices <- read_csv("data/housing_prices.csv")
```

```{r}
glimpse(housing)
```

```{r}
summary(housing)
```

# 2. 
We expect the total_rooms of houses to be strongly correlated with total_bedrooms. Use ggpairs() to investigate correlations between these two variables.

```{r}
housing %>% 
  select(total_rooms, total_bedrooms) %>% 
  ggpairs(progress = FALSE)

```

# 3. 
So, we do find significant correlations. Let’s drop total_bedrooms from the dataset, and use only total_rooms going forward.

```{r}
housing_trim <- housing %>% 
  select(-total_bedrooms)

housing_prices_trim <- housing_prices %>% 
  select(-total_bedrooms)
```

#### Added At Review -----------------------------------------------------------

We could engineer some new variables.

```{r}
housing_trim_eng <- housing_trim %>%
  mutate(rooms_per_house = total_rooms / households,
         people_per_house = population / households,
         rooms_per_person = total_rooms / population)

housing_trim_eng

housing_prices_eng <- housing_prices_trim %>%
  mutate(rooms_per_house = total_rooms / households,
         people_per_house = population / households,
         rooms_per_person = total_rooms / population)

housing_prices_eng


```

And take a look at these using `skim()` - which when we do, we can see that a lot of the variables are positively skewed.

Positively skewed data is not good for linear models.

```{r}

housing_trim_eng %>% 
  skimr::skim() %>% 
  view()

```

Where we have skewed data that is non-negative, we can log transform it to make it more normal.

```{r}

housing_prices_eng %>%
  ggplot(aes(x = median_income)) +
  geom_histogram()

housing_prices_eng %>%
  ggplot(aes(x = median_house_value)) +
  geom_histogram()

```

```{r}
housing_log <- housing_prices_eng %>% 
  select(-c(ocean_proximity, latitude, longitude, housing_median_age)) %>% 
  mutate(across(everything(), log)) %>% 
  rename_with(~ paste0("log_", .x)) %>% 
  bind_cols(housing_prices_eng)
```

```{r}
housing_log %>%
  ggplot(aes(x = log_median_income)) +
  geom_histogram()

housing_log %>%
  ggplot(aes(x = log_median_house_value)) +
  geom_histogram()
```

With values now transformed and more normal. Lets have a look for some relationships.

```{r}
housing_log %>% 
  ggplot(aes(x = median_income, y = median_house_value)) +
  geom_point()
```

```{r}

housing_log %>% 
  ggplot(aes(x = ocean_proximity, y = median_house_value)) +
  geom_boxplot()

```

There appears to be some kind of relationship between ocean proximity and house value.

```{r}
housing_log %>% 
  ggplot(aes(x = longitude, y = latitude, colour = ocean_proximity)) +
  geom_point()
```

Lets group some of those levels!

```{r}
housing_ocean <- housing_log %>% 
  mutate(ocean_prox_grouped = if_else(
    ocean_proximity %in% c("<1H OCEAN", "NEAR BAY", "NEAR OCEAN"), "NEAR WATER", ocean_proximity))

housing_ocean
```

```{r}
housing_ocean %>% 
  ggplot(aes(x = longitude, y = latitude, colour = ocean_prox_grouped)) +
  geom_point()
```

```{r}

housing_ocean %>%
  select(log_median_house_value, 
         log_total_rooms, 
         log_population, 
         log_households,
         log_rooms_per_house,
         log_people_per_house,
         log_rooms_per_person) %>% 
  ggpairs(progress = FALSE)

```

```{r}
housing_ocean %>%
  select(log_median_house_value, housing_median_age, log_median_income, ocean_prox_grouped) %>% 
  ggpairs(aes(colour = ocean_prox_grouped, 
              alpha = 0.5),
          progress = FALSE)
```

```{r}
model1 <- lm(log_median_house_value ~ log_median_income,
             data = housing_ocean)

autoplot(model1)
```

```{r}
summary(model1)
```


Once have a Simple Linear Regression we want to identify which predictors to add next.

We do this by considering our residuals.

```{r}
library(modelr)
```

```{r}
housing_ocean %>% 
  add_residuals(model = model1) %>%  # Adds redis column, difference from actual to fitted median_house_value 
  select(log_median_house_value, 
         ocean_prox_grouped, 
         log_rooms_per_person, 
         log_people_per_house, 
         resid) %>% 
  ggpairs(progress = FALSE)
```


```{r}
model1 <- lm(log_median_house_value ~ log_median_income,
             data = housing_ocean)

autoplot(model1)
```

# 4.1

We are interested in developing a regression model for the median_house_value of a house in terms of the possible predictor variables in the dataset.

Use ggpairs() to investigate correlations between median_house_value and the predictors (this may take a while to run, don’t worry, make coffee or something).

```{r}
ggpairs(housing_trim, progress = FALSE)
```

# 4.2
Perform further ggplot visualisations of any significant correlations you find.

`ggpairs()` shows a strong correlation between `median_house_value` and `median_income`, with a Correlation Coefficient of 0.688. 

This is visualised below.

```{r}

housing_trim %>% 
  ggplot(aes(x = median_income, y = median_house_value)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

```

# 5. 
Shortly we may try a regression model to fit the categorical predictor ocean_proximity. Investigate the level of ocean_proximity predictors. How many dummy variables do you expect to get from it?

There are 5 levels of `ocean_proximity` so we would use 4 dummies. 

```{r}

housing_trim %>% 
  distinct(ocean_proximity)
```

# 6.
Start with simple linear regression. Regress median_house_value on median_income and check the regression diagnostics.

```{r}
housing_model <- lm(median_house_value ~ median_income, 
                    data = housing_trim)
```

```{r}
autoplot(housing_model)
```

Plot 1: I'm unsure of the correct interpretation of this; however, the smoothed line suggests no clear pattern to distribution staying relatively flat and close to 0. To the my eye though it looks as though lower fitted values are over-estimated and higher fitted values are under-estimated. 
Plot 2: There is some variation from the normal distribution, although the residuals are relatively normally distributed.
Plot 3: I'm unsure on the correct interpretation of this; however, the smoothed line stays relatively constant/flat although to my eye there looks to be funneling at the higher end of the fitted values. 

# 7
Add another predictor of your choice. Check your assumptions, diagnostics, and interpret the model.

```{r}
housing_trim %>% 
  ggplot(aes(x = median_house_value, y = ocean_proximity))+
  geom_boxplot()
```

```{r}
housing_model_2 <- lm(median_house_value ~ median_income + ocean_proximity, 
                    data = housing_trim)
```

```{r}
autoplot(housing_model_2)
```

My interpretations here are the same as the first model, although again I'm a little unsure if these are correct.

Plot 1: I'm unsure of the correct interpretation of this; however, the smoothed line suggests no clear pattern to distribution staying relatively flat and close to 0. To the my eye though it looks as though lower fitted values are over-estimated and higher fitted values are under-estimated. 
Plot 2: There is some variation from the normal distribution, although the residuals are relatively normally distributed.
Plot 3: I'm unsure on the correct interpretation of this; however, the smoothed line stays relatively constant/flat although to my eye there looks to be funneling at the higher end of the fitted values. 

```{r}
summary(housing_model_2)
```

- The Adjusted R-squared value is 0.5879 meaning that together the predictors of median_income and ocean_proximity explain 69% of median_house_value variation.

- The p-values for each of the individual predictors are very low, suggesting that all predictors statistically significant.


