---
title: "Week 2 / Day 1 / Homework"
output: html_notebook
---

# Load Libraries
```{r}
library(tidyverse)
library(janitor)
```

# MVP

## 1.1 Joins

### Q1. 
Read in all 4 credit card transaction datasets and clean column names.

#### Read In Data
```{r}
card_data <- read_csv("data/CardBase.csv")
customer_data <- read_csv("data/CustomerBase.csv")
fraud_data <- read_csv("data/FraudBase.csv")
transaction_data <- read_csv("data/TransactionBase.csv")

# reading in the data and cleaning the names could've been done in one step:

# card_data <- read_csv("data/CardBase.csv") %>% clean_names()
```

#### Explore Data
```{r}
glimpse(card_data)
```

```{r}
glimpse(customer_data)
```

```{r}
glimpse(fraud_data)
```

```{r}
glimpse(transaction_data)
```

#### Clean column names

```{r}
card_data <- card_data %>% 
  clean_names()
```

```{r}
customer_data <- customer_data %>% 
  clean_names()
```

```{r}
fraud_data <- fraud_data %>% 
  clean_names()
```

```{r}
transaction_data <- transaction_data %>% 
  clean_names()
```

### Q2.
Join the data containing card details and customer details by customer id, so that all records of card details and any matching records in customer details are kept.

```{r}
card_customer_data <- left_join(card_data, customer_data, "cust_id")
```

### Q3.
Join the data containing fraud details with transaction details so all rows of both tables are kept. What does the resulting row number tell you?

```{r}
fraud_transaction_data <- full_join(fraud_data, transaction_data, "transaction_id")
```

The resulting number of rows is 10,000. This is the same as the original `transaction_data` data and tells us that ever observation in `fraud_data` matches which an observation from `transaction_data`. If there had been an observation from `fraud_data` that had not matched with an observation from `transaction_data`, because a `full_join` is used, this would have added additional rows and the final row count would have been greater than 10,000.

### Q4.
Join the data containing card details with transaction details so rows from the first which have matching ones in the second are returned, but only return rows in the first table once.

```{r}
card_transaction_data <- semi_join(card_data, transaction_data, 
                                   by = c("card_number" = "credit_card_id"))

```

The resulting number of rows is 500. This is the same as the original `card_data` data and tells us that evert card from this data was found in the `transaction_data` data - meaning every card has made at least one transaction.

## 1.2 Tidy Data

### Q5.
Read in hat_observations and separate observation into two columns, hat_colour and hat_type.

```{r}
hat_data <- read_csv("data/hat_observations.csv")

```

```{r}
hat_data_seperated <- hat_data %>% 
  separate(observation, into = c("hat_colour", "hat_type"), ",")

```

### Q6.
Unite day, month, and year columns into a column called date using a suitable separator.

```{r}
# Unite the year, month and day columns into one `date` column separated by a "/"

hat_data_united <- hat_data_seperated %>% 
  unite(date, c("year", "month", "day"), sep = "/")
```

```{r}
# Format `date` column using lubridate, to ensure consistent formatting of day and month values by updating all days and months to two digits eg. update `1` to `01`.

hat_data_tidy <- hat_data_united %>%
  mutate(date = ymd(date))
```

Then find the date where the most berets were observed.

```{r}
hat_data_tidy %>%
  filter(hat_type == "beret") %>% 
  group_by(date) %>% 
  summarise(total_observations = sum(observation_count)) %>%
  slice_max(total_observations, n = 1)

```

# Extension

## 2.1 Joins

### Q1.
Can you join all 4 datasets together so that you’re left with a dataset that looks like below with 109 rows and 12 columns?

```{r}
# Join `transaction_data` where there is a match to `card_customer_data`.

card_customer_transaction_data <- left_join(card_customer_data, transaction_data,
                                            by = c("card_number" = "credit_card_id"))

```

```{r}
# Filter `card_customer_transaction_data` leaving only entries which have a match in `fraud_data`.

card_extension <- semi_join(card_customer_transaction_data, fraud_data,
                                 by = "transaction_id")

```

```{r}
# Add the `fraud_flag` column from `fraud_data` where their is a match in `card_extension`.

card_extension_data <- left_join(card_extension, fraud_data,
                                 by = "transaction_id")

```

## 2.1 tidyr

### Q2.
Read in exam_scores and transform it into long format with two new columns exam_question and score. 

```{r}
exam_data <- read_csv("data/exam_scores.csv")
```

```{r}
glimpse(exam_data)
```

```{r}
# Pivot the data to long format.

exam_data_long <- exam_data %>% 
  pivot_longer(cols = exam_Q1:exam_Q10,
               names_to = "exam_question",
               values_to = "score"
  )

```

Then, using separate and select, remove superfluous information from the values in exam_question

```{r}
# Use separate to separate the question number from leading text, then use select to drop the column created containing the unecessary leading text.

exam_data_seperated <- exam_data_long %>% 
  separate(exam_question, into = c("superfluous_information", "question"), "Q") %>% 
  select(-"superfluous_information")

```


