---
title: "Week 2 / Day 3 / Homework"
output: html_notebook
---

## MVP

```{r}
library(tidyverse)
```


### Q1.
Find the weather forecast for the next three days online. Create a list of lists in which each inner list should be labelled with the name of the day of the week. Each inner list should contain named elements telling you the temperature, chance of rain and wind speed.

```{r}

weather <- list(
  Thursday_30 = list(max_temp = 14, rain_at_noon = 16, wind_speed_at_noon = 14),
  Friday_01 = list(max_temp = 14, rain_at_noon = 10, wind_speed_at_noon = 13),
  Saturday_02 = list(max_temp = 11, rain_at_noon = 22, wind_speed_at_noon = 13)
)

weather

```

### Q2.
Using the list you’ve just made find:

The chance of rain tomorrow

```{r}

weather$Thursday_30$rain_at_noon

```

The wind speed on the day after tomorrow

```{r}

weather$Friday_01$wind_speed_at_noon

```

The temperature tomorrow.

```{r}

weather$Thursday_30$max_temp

```

### Q3.
Find a 14 day forecast online and create a matrix of the forecasts for temperature, where each week is one row and columns correspond to days. You will want to give a name to the matrix you create.

```{r}

days <- c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")
week_num <- c("Week 1", "Week 2")
temps <- c(17, 19, 20, 14, 12, 16, 17, 21, 4, 7, 10, 14, 17, 20)

weather_14_days <- matrix(temps, nrow = 2)

colnames(weather_14_days) <- days
rownames(weather_14_days) <- week_num

weather_14_days

```

### Q4.
From the matrix you just made select:

The third day of the second week

```{r}

weather_14_days[2, 3]

```

The whole second week

```{r}

weather_14_days[2,]

```

The first, third and fifth days in both weeks.

```{r}

weather_14_days[, c(1, 3, 5)]

```

### Q5.
Write a function that pre-appends “The answer is:” to any character input, while for numeric input, it either just returns the number for numbers less than 10, or divides the number by 10 for numbers greater than 10. For anything else return “Invalid input”.

```{r}
my_function <- function(x) {
  
  if (is.character(x)) { 
    str_c("The answer is ", x)
  } 
  else if (is.numeric(x) & (x < 10)) {
    print(x)
  } 
  else if (is.numeric(x) & (x > 10)) {
    print(x / 10)
  } 
  else {
    print("Invalid Input")
  }
}    
  
```

```{r}
my_function(input1) # Working
my_function(input2) # Working
my_function(input3) # Working
my_function(input4)  # Working
```

### Q5. UPDATED FROM REVIEW

```{r}

input1 <- "5"
input2 <- 9
input3 <- 11
input4 <- 10

format_answer <- function(x) {
# CHECK: if the input is a character. 
if (is.character(x)) {
# DO THIS: is the input is a character.
  print(str_c("The answer is: ", x))
# CHECK: If the input has been checked and it isn't a character, check if it's a number. 
} else if (is.numeric(x)) {
# CHECK: If the input is a number check if it's less than 10. 
  if (x < 10) {
# DO THIS: if the number is less than 10.
    print(x)
# CHECK: If the number is not less than 10, check if it is greater than 10.
  } else if (x > 10) {
# DO THIS: if the number if greater than 10.
    print(x / 10)
# DO THIS: If the number is not greater or less than ten (ie. it is 10) do this.
  } else {
    print("Invalid Input")
  }
# DO THIS: if the input is not a number or a character.
} else {
  print("Invalid Input")
  }
}

```

```{r}

format_answer(input1)
format_answer(input2)
format_answer(input3)
format_answer(input4) 

```

### Q6.
Write a function which checks if two vectors have the same elements.

If they have the same elements in the same order, return “Exactly the same”
If they have the same elements but in a different order, return “The same”
If they don’t have the same elements, return “Different”.

```{r}

vector_1 <- c(1, 2, 3)
vector_2 <- c(1, 2, 3)
vector_3 <- c(3, 2, 1)
vector_4 <- c(5, 10, 15)

```

```{r}

my_function_1 <- function(a, b) {
  
  if (all(a == b)) { 
        print("Exactly the same")
  } else if (all(sort(a) == sort(b))) {
        print("The same")
  } else if (all(sort(a) != sort(b))) {
        print("Different")
  }
}   

```

```{r}
my_function_1(vector_1, vector_2) # Should return "Exactly the same"
my_function_1(vector_1, vector_3) # Should return "The same"
my_function_1(vector_1, vector_4) # Should return "Different"
```

### Q6. UPDATED FROM REVIEW

```{r}
# In the below version, we use an else rather than a final else if logical check.

my_function_2 <- function(a, b) {
  
  if (all(a == b)) { 
        print("Exactly the same")
  } else if (all(sort(a) == sort(b))) {
        print("The same")
  } else {
        print("Different")
  }
} 
```

```{r}
my_function_2(vector_1, vector_2) # Should return "Exactly the same"
my_function_2(vector_1, vector_3) # Should return "The same"
my_function_2(vector_1, vector_4) # Should return "Different"
```

