---
title: "Week 10 / Day 3 / Review"
output: html_notebook
---

# Load Libraries / Read Data

```{r}
library(tidyverse)
library(modelr)
library(GGally)
library(ggfortify)
library(skimr)
```

```{r}
kc_houses <- read_csv("data/kc_house_data.csv")
```

# Explore Data

We want to build an explanatory model for `price`, so lets look at its distribution. We can see from the below that it's highly right skewed.

```{r}
ggplot(data = kc_houses, aes(x = price)) +
  geom_histogram()
```

## Feature Engineering

```{r}
kc_houses_trim <- kc_houses %>%
  select(-id, -date) %>% 
  mutate(has_been_renovated = yr_renovated != 0,
         waterfront = as.logical(waterfront),
         grade = as.factor(grade)) %>% 
  select(-yr_renovated)
```

Rather than running `ggpairs()` across all variables, try grouping these into variables you think are associated.

### Explore Group 1

Group 1: Size of House: bedrooms, bathrooms, sqft_living, sqft_lot, floors, sqft_above, sqft_above

```{r}
kc_houses_trim %>%
  select(price, bedrooms, bathrooms, sqft_living, sqft_lot, floors, sqft_above, sqft_above) %>% 
  ggpairs(progress = FALSE)
```

### Explore Group 2

Group 2: Location: waterfront, view, lat, long, zipcode

```{r}
kc_houses_trim %>%
  select(price, waterfront, view, lat, long, zipcode) %>% 
  ggpairs(progress = FALSE)
```

### Explore Group 3

Group 3: Current State: has_been_renovated, condition, grade

```{r}
kc_houses_trim %>%
  select(price, has_been_renovated, condition, grade) %>% 
  ggpairs(progress = FALSE)
```

### Explore Group 4

Group 4: Age: yr_built

```{r}
ggplot(kc_houses_trim, aes(x = yr_built, y = price)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", colour = "red")
```

## Select Initial Predictor Candidates from Each Group

After exploration, the following look like to candidates:

- sqft_living
- lat or waterfront
- grade
- yr_build

## Check for Correlation & Aliases In Selected Initial Predictor Candidates

```{r}

kc_houses_trim %>% 
  select(price, sqft_living, lat, waterfront, grade, yr_built) %>% 
  ggpairs(progress = FALSE)

```

```{r}
alias(price, sqft_living, lat, watefront, grade, yr_built, 
      data = kc_houses_trim)
```

## Create Candidates Tibble

```{r}
kc_houses_candidates <- kc_houses_trim %>% 
  select(price, sqft_living, lat, waterfront, grade, yr_built)
```

# Build Model

```{r}
model_1 <- lm(log(price) ~ sqft_living, data = kc_houses_candidates)
```

## Check Model Summary

```{r}
summary(model_1)
```

## Check Diagnostics

```{r}
autoplot(model_1)
```

# 2nd Predictor

## Add Residuals

Remember to remove predictors that are already in the model - `sqft_living` and `price`.

```{r}

kc_resid <- kc_houses_candidates %>%
  add_residuals(model_1) %>% 
  select(-sqft_living, -price)

```

## Check Remaining Predictor Candidates Against Residuals for Correlation

From the below we can see that there appears to be a strong correlation between `lat`, `waterfront` and `grade` and the residuals.

```{r}
kc_resid %>% 
  ggpairs(progress = FALSE)
```
## Create 2nd Model

```{r}
model_2a <- lm(log(price) ~ sqft_living + lat, data = kc_houses_candidates)
```

## Check Model Summary

```{r}
summary(model_2a)
```

## Check Diagnostics

```{r}
autoplot(model_2a)
```

While the diagnostics show an increase in heteroskedasticity, the residuals appear to show no clear trend and are normally distributed.

Final Model: log(price) ~ sqft_living + lat

Interpretation: The price of a house in KC country can be explained by a combination of it's size (sqft of the living space) and it's latitude (Northerness).

Example: For a house at 1910 sqft (median sqft) located at latitude 47.56 (median lat), the predicted price would be $437884.3 (low estimate = 321261.3, high estimate = 596843.4).


as calculated below.

Step 1.
outcome = b0 + b1 * 1st Predictor + b2 * 2nd Predictor

Step 2.
log(price) = b0(intercept) + b1(coefficient) * 1st Predictor + b2(coefficient) * 2nd Predictor

Step 3.
log(price) = b0(intercept) + b1(coefficient) * `sqft_living` + b2(coefficient) * `lat`

Step 4.
log(price) = -62.56 + 3.863e-04 * `1910` + 1.573 * `47.56` = 12.98971

Step 5.
price = e^12.98971 (exp 12.98971 ) = 437884.3

```{r}

# Taking log(price) and RSE, we can say price within +- bracket.

exp(12.98971 + c(-0.3097, 0.3097))

```



