---
title: "Week 5, Day 1, Homework"
output: html_notebook
---

# Load Libraries

```{r}
library(tidyverse)
library(janitor)
library(scales)
```

# Read In Data

```{r}

qb_revenue_breakdown <- clean_names(CodeClanData::qb_revenue_breakdown)
qb_monthly_sales <- clean_names(CodeClanData::qb_monthly_sales)
qb_competitors <- clean_names(CodeClanData::qb_competitors)
qb_device_data <- CodeClanData::qb_device_data

```

# MVP

## Part 1

### Q1.
Take the data in the dataset qb_revenue_breakdown and make a stacked bar chart showing the sources of revenue across the two years in the dataset.

```{r}

qb_revenue_breakdown %>% 
  ggplot(aes(x = year, y = revenue, fill = product)) +
  geom_col() +
    labs(
    x = "\nYear",
    y = "Revenue\n",
    title = "Revenue Sources",
    fill = "Product:"
  ) +
  theme(title = element_text(face = "bold"),
        plot.title = element_text(size = 20))

```

### Q2.
Make a line chart showing monthly sales of the “flagship product” - the ‘KwikBit Unit 1’ - over the last year. This data is inside qb_monthly_sales, but you will need to use filter() or subset() (base R).

```{r}

qb_monthly_sales %>%
  filter(cashflow == "Kwikbit Unit 1 Sales") %>% 
  ggplot(aes(x = date, y = amount)) +
  geom_line() +
  scale_x_date(date_labels = "%b %y", date_breaks = "1 month") +
  labs(
    x = "\nMonth / Year",
    y = "Sales Amount\n",
    title = "Kwikbit Unit 1 Sales In 2018 & 2019",
    fill = "Product:"
  ) +
  theme(title = element_text(face = "bold"),
        plot.title = element_text(size = 15))

```

### Q3.
Make a line chart showing monthly revenue and costs over the last year. This data is also in qb_monthly_sales.

```{r}

qb_monthly_sales %>%
  filter(cashflow != "Kwikbit Unit 1 Sales") %>% 
  ggplot(aes(x = date, y = amount, colour = cashflow )) +
  geom_line() +
  scale_x_date(date_labels = "%b %y", date_breaks = "1 month") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5),
        title = element_text(face = "bold"),
        plot.title = element_text(size = 15),
        legend.title = element_blank()) +
  labs(
    x = "\nMonth / Year",
    y = "Amount\n",
    title = "Total Revenue vs. Costs",
  ) +
  scale_color_discrete(labels=c('Costs', 'Total Revenue'))

```

### Q4.
Show annual sales of personal fitness trackers over the last 5 years broken down by company as a ribbon plot (use geom_area). This data is in qb_competitors.

```{r}

qb_competitors %>% 
  ggplot(aes(x = year, y = revenue, fill = company)) +
  geom_area(colour = "black") +
  scale_y_continuous(labels = comma) +
  theme(title = element_text(face = "bold"),
        plot.title = element_text(size = 12.5)) +
  labs(
    x = "\nYear",
    y = "Revenue\n",
    title = "Annual Personal Fitness Track Sales by Company",
    fill = "Company:"
  ) 

```

### Q5.
Now show the sales from the four competitors as a line graph. Include an extra layer that shows the data points used to make the lines.

```{r}
qb_competitors %>%
  ggplot(aes(x = year, y = revenue, colour = company)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(labels = comma) +
  theme(title = element_text(face = "bold"),
        plot.title = element_text(size = 12.5)) +
  labs(
    x = "\nYear",
    y = "Revenue\n",
    title = "Annual Personal Fitness Track Sales by Company",
    colour = "Company:"
  ) 

  
```

### Q6.
Now the company wants to compare the number of steps that their device counts vs. their competitors. Make a line graph of the number of steps throughout time, and use faceting to compare between companies and people. The data for this is in qb_device_data.

```{r}

qb_device_data_updated <- qb_device_data %>%
  group_by(id) %>% 
  mutate(rolling_steps_count = cumsum(counts),
         total_time_mins = (hours * 60) + mins)

```

```{r}

qb_device_data_updated %>% 
  ggplot(aes(x = total_time_mins, y = rolling_steps_count, colour = id)) +
  geom_line()

```


## Part 2

Take the plots that you produced in part one and now polish them by:

1. Adding appropriate labels
2. hanging the scales and coordinates when appropriate.
3. Applying a unified theme, which is described below:

- Graphs should have white backgrounds, and use colour sparingly.
- There should be faint grid lines.
- Font sizes should be ~12pt although titles should be slightly larger and axis labels can be slightly smaller.
- All plots should use colours from the following company colour scheme.

### Q1 Revised

```{r}
col_scheme <- c("#E89FE9", "#50434F", "#B6A7B5", "#F9A472", "#BD7040")
```


```{r}

qb_revenue_breakdown %>% 
  ggplot(aes(x = year, y = revenue, fill = product)) +
  geom_col() +
  theme_light() +
  labs(
    x = "\nYear",
    y = "Revenue\n",
    title = "Revenue Sources",
    fill = "Product:"
  ) +
  theme(plot.title = element_text(size = 15, face = "bold"),
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        legend.title = element_text(size = 10)) +
  scale_fill_manual(values = col_scheme)

```

### Q2 Revised

```{r}
qb_monthly_sales %>%
  filter(cashflow == "Kwikbit Unit 1 Sales") %>% 
  ggplot(aes(x = date, y = amount)) +
  geom_line(colour = "#F9A472") +
  theme_light() +
  scale_x_date(date_labels = "%b %y", date_breaks = "1 month") +
  labs(
    x = "\nMonth / Year",
    y = "Sales Amount\n",
    title = "Kwikbit Unit 1 Sales In 2018 & 2019",
    fill = "Product:"
  ) +
  theme(plot.title = element_text(size = 15, face = "bold"),
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        axis.text.x = element_text(angle = 45, vjust = 0.5))
```

### Q3 Revised

```{r}
qb_monthly_sales %>%
  filter(cashflow != "Kwikbit Unit 1 Sales") %>% 
  ggplot(aes(x = date, y = amount, colour = cashflow)) +
  geom_line() +
  scale_x_date(date_labels = "%b %y", date_breaks = "1 month") +
  theme_light() +
  theme(plot.title = element_text(size = 15, face = "bold"),
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  labs(
    x = "\nMonth / Year",
    y = "Amount\n",
    title = "Total Revenue vs. Costs",
  ) +
  scale_color_manual(values=c("#E89FE9", "#F9A472"))
```

### Q4 Revised

```{r}
qb_competitors %>% 
  ggplot(aes(x = year, y = revenue, fill = company)) +
  geom_area() +
  theme_light() +
  scale_y_continuous(labels = comma) +
  theme(plot.title = element_text(size = 13, face = "bold"),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8)) +
  labs(
    x = "\nYear",
    y = "Revenue\n",
    title = "Annual Personal Fitness Tracker Revenue by Company",
    fill = "Company:"
  ) +
  scale_fill_manual(values = col_scheme)
```

### Q5 Revised

```{r}
qb_competitors %>%
  ggplot(aes(x = year, y = revenue, colour = company)) +
  geom_line() +
  geom_point() +
  theme_light() +
  scale_y_continuous(labels = comma) +
  theme(plot.title = element_text(size = 13, face = "bold"),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8)) +
  labs(
    x = "\nYear",
    y = "Revenue\n",
    title = "Annual Personal Fitness Track Sales by Company",
    colour = "Company:"
  ) +
    scale_color_manual(values = col_scheme)
```

### Q6 Revised