---
title: "Week 6 / Day 1"
output: html_notebook
---

```{r}
# Load Libraries
library(gtools)
library(tidyverse)

# Source Probability Functions
source("prob.R")

# Read In Data and Add Logial Type `crew` Column.
titanic <- read_csv("data/titanic.csv")
titanic_with_crew <- titanic %>%
  mutate(crew = if_else(class %in% c("engineering crew", "victualling crew", 
                                     "restaurant staff", "deck crew"),
                        TRUE, FALSE))
```

## MVP

### 1.1

Calculate: prob(survived∣crew)

p(A | B) = p(A AND B) / p(B)

Event A: Survived = yes
Event B: Crew = TRUE

```{r}

# prob(survived∣crew)

titanic_with_crew %>%
  # Filters by event B.
  filter(crew == TRUE) %>% 
  # Sum where event A occurred and divide by filtered rows.
  summarise(p = sum(survived == "yes") / n())

```

Calculate: prob(survived∣passenger)

```{r}

# prob(survived∣passenger)

titanic_with_crew %>%
  # Filters by event B.
  filter(crew == FALSE) %>% 
  # Sum where event A occurred and divide by filtered rows.
  summarise(p = sum(survived == "yes") / n())

```

Compare: prob(survived∣crew) & prob(survived∣passenger) 

prob(survived∣crew) = 0.2370787
prob(survived∣passenger) = 0.3796507

The above suggests that around 24% of crew survived compared to around 38% of passengers. Therefore, it appears that passengers had a greater probability of surviving than crew members.

### 1.2 / Q1.
If I roll a six-sided dice once, what is the probability of getting a number less than 5?

Step 1.
p(event) = n(outcomes in event) / n(total outcomes in sample space)

Step 2.
p(less than 5) = {1, 2, 3, 4} / {1, 2, 3, 4, 5, 6}

Step 3.
p(less than 5) = 4 / 6 = 0.67

### 1.2 / Q2.
How many outcomes are there in the sample space of an experiment in which we toss three dice?

```{r}
s_three_dice <- rolldie(3)

s_three_dice

# Answer using `rolldie()` function: 216
```

```{r}

dice_values <- c(1, 2, 3, 4, 5, 6)

possible_permutations <- as_tibble(
  permutations(n = length(dice_values),
               r = 3,
               v = dice_values,
               repeats.allowed = TRUE))

possible_permutations

# Answer using `permutations()` function: 216

```

Answer: 216

### 1.2 / Q3.
In an experiment where we roll four dice:

a) How many outcomes are in the event where the total of the four dice equals 5, 6 or 7?

```{r}

s_four_dice <- rolldie(4)

# 1296 permutations

s_four_dice_totals <- s_four_dice %>% 
  mutate(total = rowSums(.)) %>% 
  filter(total %in% c(5, 6, 7))

```

Answer: 34

b) What is the probability of this event?

```{r}

# p(event) = n(outcomes in event) / n(total outcomes in sample space)

p <- 34 / 1296

```

Answer: 0.026

c) What assumptions do we apply in answering part ii?

Answer: That the dice are fair and shaken prior to rolling.

### 1.2 / Q4.
Five chairs are placed in a row in a room, along with five people of differing height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?

Step 1:
p(total) = p(event) + p(compliment of event) 

Step 2:
p(Decreasing Height Left To Right) = 1 - p(compliment of event)

Step 3:
p(Decreasing Height Left To Right) = 1 - 0.99

Answer: p(Decreasing Height Left To Right) = 0.01

```{r}

height_values <- c(5, 6, 7, 8, 9)

possible_height_permutations <- as_tibble(
  permutations(n = length(height_values),
               r = 5,
               v = height_values,
               repeats.allowed = FALSE))

# p(event) = n(outcomes in event) / n(total outcomes in sample space)
p_compliment <- 119 / 120

```

