---
title: "Week 6, Day 3"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(infer)
```

## MVP

### Q1.
Load the data again, clean_names(), and re-familiarise yourself with it.

```{r}

ames <- read_csv("data/ames.csv") %>% clean_names()

```

```{r}

glimpse(ames)

```

### Q2.
Investigate the distribution of lot_area. Is the distribution roughly normal? If not, what problems do you find?

The distribution is not roughly normal as there is a significant peak and there also look to be right-sided outliers suggesting it is positively skewed.

```{r}

ames %>% 
  ggplot() +
  geom_histogram(aes(x = lot_area), bins = 27)

```

### Q3.
Compute and visualise a bootstrap sampling distribution for the mean(lot_area) of the sold houses.

```{r}

ames_infer_resample_2930 <- ames %>%
  specify(response = lot_area) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "mean")

ames_infer_resample_2930

```

```{r}

ames_infer_resample_2930 %>% 
  ggplot() +
  geom_histogram(aes(x = stat), bins = 25)

```

### Q4.
Use your bootstrap distribution to calculate a 95% CI for mean(lot_area), and visualise it on the distribution.

```{r}
infer_ci_95 <- ames_infer_resample_2930 %>%
  get_confidence_interval(level = 0.95, type = "percentile")

infer_ci_95
```

```{r}

ames_infer_resample_2930 %>%
  visualise(bins = 25) +
  shade_confidence_interval(endpoints = infer_ci_95)

```

### Q5.
You would like to know the mean(lot_area) of the sold houses with higher confidence. Calculate the 99% CI for this variable (you can re-use your bootstrap distribution from above). Is it narrower or broader than the 95% CI? Does that make sense?

It is broader than 95%CI as we need to allow for more possible values in order to increase our confidence.

```{r}

infer_ci_99 <- ames_infer_resample_2930 %>%
  get_confidence_interval(level = 0.99, type = "percentile")

infer_ci_99

```

```{r}

ames_infer_resample_2930 %>%
  visualise(bins = 25) +
  shade_confidence_interval(endpoints = infer_ci_99)

```

### Q6.
Calculate the point estimate of the mean(lot_area)

The point estimate of the mean(lot_area) is 10153.5. In this case if we take our original data `ames` to be the population and calculate the mean(lot_area) we get a value of 10147.92. Here we can see that although these values are slightly different, our point estimate would have fallen within our 95% CI.

```{r}

mean_infer <- ames_infer_resample_2930 %>% 
  summarise(mean = mean(stat))

mean_infer

```

```{r}

ames %>% 
  summarise(mean = mean(lot_area))

```

